%%% !TEX program = lualatex
\documentclass[10pt,english]{article}
\usepackage[a4paper,bindingoffset=0in,left=1.3in,right=1.3in,top=1.2in,bottom=1in,footskip=.25in]{geometry}
\usepackage{lineno,hyperref}
\usepackage[margin=1cm]{caption}
% Packages I added:
%\usepackage{caption,subcaption}
% \usepackage{tablefootnote}
\usepackage{placeins} % For floatbarrier
% \usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{colortbl}
% \usepackage{arydshln} % For dashed line in table (appendix)
% \usepackage{multirow}
\usepackage{tabularx}
% \newcolumntype{R}{>{\raggedleft\arraybackslash}X}%
% \newcolumntype{C}{>{\centering\arraybackslash}X}%
% \newcolumntype{s}{>{\centering\arraybackslash\hsize=0.8\hsize}X}
% \newcolumntype{z}{>{\centering\arraybackslash\hsize=1.3\hsize}X}
% \newcolumntype{a}{>{\centering\arraybackslash\hsize=0.6\hsize}X}
% \newcolumntype{b}{>{\centering\arraybackslash\hsize=1.1\hsize}X}
% \newcolumntype{y}{>{\centering\arraybackslash\hsize=1.6\hsize}X}
% \newcolumntype{o}{>{\raggedright\arraybackslash\hsize=1.6\hsize}X}
% \newcolumntype{d}{>{\centering\arraybackslash\hsize=0.85\hsize}X}
% \newcolumntype{e}{>{\raggedright\arraybackslash\hsize=0.4\hsize}X}
% \newcolumntype{t}{>{\raggedright\arraybackslash\hsize=1.6\hsize}X}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[caption=false,font=normalsize]{subfig}
%\captionsetup[subfigure]{margin=20pt}
\usepackage{notoccite}
\usepackage{longtable}
% \usepackage{ltxtable}

\usepackage{eurosym}
\usepackage{tikz, pgfplots, graphicx}
\usetikzlibrary{spy,calc}
\usetikzlibrary{positioning}
\pgfplotsset{compat=newest}
\usepackage[space]{grffile} % For defining path variables with spaces
\usepgfplotslibrary{external} 
% \tikzexternalize[prefix=tikz/] % automatically saves pdfs generated by tikz into a folder tikz



\input{00_math_symbols}


% \bibliographystyle{elsarticle-num}



% \usepackage[colorlinks]{hyperref}
% \usepackage{breakurl}

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor ESOMs}


% \allowdisplaybreaks

% \bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

%\definecolor{LightCyan}{rgb}{0.88,1,1}
\interfootnotelinepenalty=10000

\newcommand{\myparagraph}[1]{\paragraph{#1}\mbox{}\\} % For having a new line after the paragraph title


\hyphenation{Euro-pean ap-pro-xi-ma-ted inte-gra-ting}
\title{Spine Model formulation and design}
\date{}
\author{Kris Poncelet\textsuperscript{*,\#}\\ 
{\small \textsuperscript{\#}dr. ir. Kris Poncelet, Leuven, Belgium, kris.poncelet@mech.kuleuven.be} \\
{\small \textsuperscript{*}University of Leuven (KU Leuven) Energy Institute TME Branch, EnergyVille}}

\begin{document}

\voffset = 0pt

\maketitle{}


\textbf{Included in this document version:}
\begin{itemize}
\item Basic definition of variables, parameters and equations in order to:
\begin{itemize}
\item Establish a set of commodity flow(s) which define the capacity of the unit/technology 
\item A constraint restriciting the commodity flow(s) to the installed capacity of the unit/technology
\item Establisch a linear relationship between different commodity flows entering and leaving a specific unit/technology (in part)
\item Detailed technical constraints (UC type of constraints)
\item A commodity balance constraint
\end{itemize} 
\end{itemize} 


\textbf{Not included:}
\begin{itemize}
\item Geographical regions (treatment of nodes) and network related constraints (Maren, Jody)
\item Storage technologies
\item Temporal model structure (Juha? + Kris)
\item Objective function + flexibility
\item Bounds on absolute or total commodity flows
\item Stochastics (Juha?)
\end{itemize} 

\textbf{Open issues}: 
\begin{itemize}
\item Exogenous commodities: commodity input units or not?
\item What needs to be adapted if different commodity flows are tracked at a different level of temporal granularity?
\item Linking of different equations 'selected" by the archetype (Kris)
\item Feasible to treat reserve capacity as a commodity? -> Different types of reserves and constraints for reserve capacity dependent on type (e.g., spinning, non-spinning, upward, downward, ...) (Kris)
\item Dealing with the ability to have operational and investment planning problems. Many equations contain capacity-related terms. However, depending on the problem, these can be variables or parameters. Need duplicates of equations for both or does Julia/Jump has features to deal with this? (Kris)
\end{itemize}

\textbf{Next steps}:
\begin{itemize}
\item Compare to Backbone implementation + discuss!
\item Experiment with different archetypes for the coal-fired power plant example
\item Introduce the CHP example
\end{itemize}



\clearpage
\tableofcontents


\clearpage
\include{00_nomenclature}







\input{Sections/Spine_model_formulation}

\input{Sections/Open_issues}

\input{Sections/Examples}




















% \bibliographystyle{abbrv}
% \bibliography{mybibfile}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \bibliographystyle{abbrv}
% \bibliography{mybibfile}

\end{document}